Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-03-03 abstract.tex
almost final introduction. i also polished the abstract

% 
This article presents a seismic hazard assessment for northern Iran where a smoothed seismicity approach has been used in combination with an updated seismic catalog and a ground motion prediction equation recently found to yield the best fit with data. We evaluate the hazard over a geographical region including the three major seismic zones affecting the northern part of the country, namely Azerbaijan, the Alborz Mountain Range, and Kopeh-Dagh. For completeness, we also consider the partial contribution of the two remaining seismic zones to the south, that is, Zagros to the southwest, and the south-central and southeastern Iranian seismic zones. Following the concepts of smoothed seismicity, events are not assigned to specific faults but are instead assumed to be potential seismogenic sources by in themselves, and are spatially distributed within regular grid cells of size equal to 0.1\textdegree{} in both the latitude and longitude directions. After performing the corresponding magnitude conversions, we decluster both historical and instrumental seismicity catalogs to obtain earthquake rates based on the number of events within each cell, and smooth the results to account for the uncertainty in the spatial distribution of future earthquakes. Smoothing was done using a Gaussian filter in incremental intervals of magnitude. Seismicity parameters are calculated and smoothed for each seismic zone separately, and for comparison purposes, we also consider the combination of all the seismic zones as a single seismotectonic region. Seismic hazard curves obtained for two distinct models defined according to two structural risk scenarios. These scenarios are based on the assumption that damage occurs above earthquake magnitude thresholds of $M_w$ greater than 4.5 for unreinforced and informal construction, and greater than 5.0 for modern structures following design regulations. We also present results for a combined scenario based on the proportion of reinforced and unreinforced structures. The results are presented in terms of expected peak ground acceleration curves and hazard maps, considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions. We compare our results to equivalent estimations available in the literature at selected locations and offer our perspective on the difference observed due to the intrinsic differences in the methods employed here in light of results obtained by other previous studies. In conclusion, we find our results to be helpful in understanding seismic hazard for norther Iran from a regional perspective, but highlight the fact that different methods may need to be combined to obtain a more robust estimates at specific areas of interest.

@@

>>>>

% 
This article presents a seismic hazard assessment for northern Iran, where a smoothed seismicity approach has been used in combination with an updated seismic catalog and a ground motion prediction equation recently found to yield the best fit with data. We evaluate the hazard over a geographical area including the seismic zones of Azerbaijan, the Alborz Mountain Range, and Kopeh-Dagh. For completeness, we also consider the partial contribution of the seismic zones of Zagros to the southwest, and the south-central and southeastern Iranian seismic zones. In the chosen approach, seismic events are not assigned to specific faults but are instead assumed to be potential seismogenic sources by in themselves, and are spatially distributed within regular grid cells of size equal to 0.1\textdegree{} in both latitude and longitude. After performing the corresponding magnitude conversions, we decluster both historical and instrumental seismicity catalogs to obtain earthquake rates based on the number of events within each cell, and smooth the results to account for the uncertainty in the spatial distribution of future earthquakes. Seismicity parameters are calculated for each seismic zone separately, and for comparison purposes we also consider the combination of all zones into a single seismotectonic region. Seismic hazard curves are obtained for two distinct models defined based on the assumption that damage occurs above earthquake magnitude thresholds of $M_w$ greater than 4.5 and 5.0 for unreinforced and reinforced structures, respectively. We also obtain results for a combined scenario of these two models. The results are presented in terms of expected peak ground acceleration curves and hazard maps, considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions. We compare our results to equivalent estimations available in the literature at selected locations and offer our perspective on the differences observed. In conclusion, we find our results to be helpful in understanding seismic hazard for northern Iran from a regional perspective, but highlight the fact that different methods may need to be combined to obtain more robust estimates at specific areas of interest.

@@

% Smoothing was done using a Gaussian filter in incremental intervals of magnitude.

% greater than 4.5 for unreinforced and informal construction buildings, and greater than 5.0 for modern structures following design regulations

### 2016-03-03 introduction.tex
almost final introduction. i also polished the abstract

This article presents a new seismic hazard assessment for northern portion of the Iranian plateau, using a smoothed seismicity approach in combination with an updated seismic catalog and a ground motion prediction equation recently evaluated and found to be the most reliable attenuation relationship to predict peak ground response in the region. The Iranian plateau is located on the Himalayan-Alpine seismic belt. It is confined between the convergent movements of the Arabian and Eurasian plates, and laterally trapped between the Arabian and eastern Asia-Minor to the west, and the Indian and Eurasian plates to the east. This entrapment lifts the plateau and gives rise to the three major mountain ranges delimiting Iran's land, the Zagros mountain ranges running from the southeast to the northwest, the Alborz ranges to the north, and the east Iranian ranges \citep[e.g.,][]{Berberian_1981_Chap}. It also defines the seismic activity of the region, one with a long history of large magnitude ($M>7$) earthquakes dating back to the eight century, which has repeatedly damage Iran's infrastructure and taken the lives of thousands.

@@

Over the last three decades there have been multiple efforts to better estimate the seismic hazard and assess seismic risk in Iran \citep[e.g.,][]{Tavakoli1999} [\textcolor{red}{Add at least two more references here}]. These studies have been, in part, intended to help improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. While improvements in design practice are reflected in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran, other indicators suggest high levels of seismic and therefore more complete and accurate assessments of seismic hazard, that incorporate the latest available data and alternative methods, are necessary.

In this article, we are particularly interested in the assessment of the seismic hazard of northern Iran. This region houses nearly 40 percent (31.5 million) of the total population of the country in the cities of Tehran, Mashhad, Tabriz, Karaj, Qom, Orumieh, and Rasht; and has suffered the devastating effects of earthquakes in the past (e.g., \textcolor{red}{two references}). Due to its importance, northern Iran has recently been the subject of various efforts to better characterize seismic hazard in the region \citep[e.g.,][]{Ghodrati2011, Abdollahzadeh2014a, Boostan2015} \textcolor{red}{[Are these good references? we need them to be about northern Iran. Any other?]}. In general, these studies highlight the significant seismic hazard of the region, and... \textcolor{red}{[need punch line conclusions here]}.

However useful, most previous studies focus on particular locations or are based on specific categorization of seismogenic faults, which are often subjectively interpreted and limited to the availability data. Here, we use the smoothed seismicity method introduced by \citet{Frankel1995}. This approach has been used with success in other regions (\textcolor{red}{[two references needed here]}). Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that probabilistic seismic hazard analysis based on fault systems can be more realistic and accurate, by choosing the smoothed seismicity method we seek to reduce the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh; but for completeness, we also consider the partial contribution of the areas of Zagros and the central-east seismic zones within our domain of interest (see Figure \ref{fig:Iran}). In addition, for comparison purposes, we obtain results for the the combination of all seismic zones as a uniform region.

>>>>

This article presents a new seismic hazard assessment for northern portion of the Iranian plateau, using a smoothed seismicity approach in combination with an updated seismic catalog and a ground motion prediction equation recently evaluated and found to be the most reliable attenuation relationship to predict peak ground response in the region.

The Iranian plateau is located on the Himalayan-Alpine seismic belt. It is confined between the convergent movements of the Arabian and Eurasian plates, and laterally trapped between the Arabian and eastern Asia-Minor to the west, and the Indian and Eurasian plates to the east \citep{Berberian_1981_Chap}. This entrapment lifts the plateau and gives rise to the three major mountain ranges delimiting Iran's land, the Zagros mountain ranges running from the southeast to the northwest, the Alborz ranges to the north, and the east Iranian ranges. It also defines the seismic activity of the region---one with a long history of large magnitude ($M>7$) earthquakes that are well documented dating back to the eight century---which has repeatedly damaged Iran's infrastructure and taken the lives of thousands.

@@

Over the last three decades there have been multiple efforts to better estimate the seismic hazard and assess seismic risk in Iran \citep[e.g.,][]{Tavakoli1999, Moinfar_2000_Chap, Ghodrati2003, Moinfar_2012_WCEE}. These studies have been, in part, intended to help improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. While improvements in design practice are reflected in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran, other indicators suggest high levels of seismic risk. Therefore, more complete and accurate assessments of seismic hazard---incorporating the latest available data and alternative methods---are necessary.

Northern Iran houses nearly 40 percent (31.5 million) of the total population of the country in the cities of Tehran, Mashhad, Tabriz, Karaj, Qom, Orumieh, and Rasht. This region has suffered the devastating effects of earthquakes in the past \citep[e.g.,][]{Mehrain_1990_Tech, Chafory-Ashtiany_1999_DPM, Razzaghi_2012_Tech}. Due to its importance, northern Iran has recently been the subject of various efforts to better characterize seismic hazard in the region \citep[e.g.,][]{Abdollahzadeh2014a, Boostan2015}. In general, these studies highlight the significant seismic hazard of the region, and... \textcolor{red}{[need punch line conclusions here]}.

However useful, most previous studies focus on particular locations or are based on specific categorization of seismogenic faults, which are often subjectively interpreted and limited to the availability data. Here, we use the smoothed seismicity method introduced by \citet{Frankel1995}. This approach has been used with success in other regions \citep[e.g.][]{Boyd2008, Kalkan2009, Moschetti2014}. Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that probabilistic seismic hazard analysis based on fault systems can be more realistic and accurate, by choosing the smoothed seismicity method we seek to reduce the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh; but for completeness, we also consider the partial contribution of the areas of Zagros and the central-east seismic zones within our domain of interest (see Figure \ref{fig:Iran}). In addition, for comparison purposes, we obtain results for the the combination of all seismic zones as a uniform region.
